---
title: "MachlearnAssignment5"
author: "Abhishek_Sau"
date: "26/04/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Hierarchical Clustering

The dataset "Cereals.csv" includes nutritional information, store display, and consumer ratings for 77 breakfast cereals. Data Preprocessing. Remove all cereals with missing values.

```{r}
setwd("D:\\Study\\Assignments\\MachLearn\\MachLearnAssignment5_Hierarchical")
cereals_df<- read.csv("Cereals.csv")
str(cereals_df)
head(cereals_df)
```


```{r}

library(DataExplorer)

introduce(cereals_df) #No. of missing values

cereals_df1<-na.omit(cereals_df) #dataset with omitted rows with missing values

```
Apply hierarchical clustering to the data using Euclidean distance to the normalized measurements.

```{r}
library(tidyverse)  
library(factoextra) 
library(dendextend)
library(cluster)
library(fastDummies)

#Identifying categorical and numeric variables


cereals_df1$name<-as.factor(cereals_df1$name)
cereals_df1$mfr<-as.factor(cereals_df1$mfr)
cereals_df1$type<-as.factor(cereals_df1$type)
cereals_df1$shelf <- as.factor(cereals_df1$shelf)

#creating dummy variables
vaar <- colnames(cereals_df1)
num_var <- c("calories", "protein", "fat", "sodium","fiber","carbo","sugars","potass","vitamins","weight","cups","rating")

cat_var<-cereals_df1[which(colnames(cereals_df1) %in% c('name','mfr','type','shelf'))]
cat_var<-data.frame(apply((cereals_df1[which(colnames(cereals_df1) %in% c('name','mfr','type','shelf'))]),2,as.factor))


dummy_vars <- fastDummies::dummy_columns(cat_var %>% select(-name))
num_vars <- cereals_df1[,c(4:12, 14:16)]
cereals_df2 <-cbind(cereals_df1$name,dummy_vars,num_vars) %>% select(-c(mfr, type, shelf))

```

Normalizing the data set

```{r}

cereals_df2[,c(2:25)]<- scale(cereals_df2[,c(2:25)], scale = TRUE, center = TRUE)

```

Q1. Use Agnes to compare the clustering from single linkage, complete linkage, average linkage, and Ward. Choose the best method.

```{r, fig.height=8, fig.width=14}
hc1 <- agnes(cereals_df2, method = "complete")
hc2 <- agnes(cereals_df2, method = "average")
hc3 <- agnes(cereals_df2, method = "single")
hc4 <- agnes(cereals_df2, method = "ward")

ac <- c(hc1$ac,hc2$ac,hc3$ac, hc4$ac)
ac_method <- c(hc1$method,hc2$method,hc3$method, hc4$method)
ac_df <- data.frame(ac_method, ac)
ac_df

pltree(hc4, cex = 0.6, hang = -1, main = "Dendrogram based on ward", labels = cereals_df2$`cereals_df1$name`)

```

As per the above table, ward method has the highest agglomerative coefficient, i.e closest to one. Hence, it gives the best clusters.


Q2. How many clusters would you choose?

```{r}

fviz_nbclust(cereals_df2, hcut, method = "wss")+
 geom_vline(xintercept = 2, linetype = 5)

fviz_nbclust(cereals_df2, hcut, method = "silhouette")

cereals_df2<- cereals_df2 %>% mutate(cluster= cutree(hc4, k=2))

```
As per the dendogram, I will choose 2 clusters.

Q3. Comment on the structure of the clusters and on their stability.

```{r}
library(caret)
set.seed(12)

split_index <- createDataPartition(cereals_df2$rating, p=0.6, times = 1, list = FALSE)

cereal_part1 <- cereals_df2[split_index, ]
cereal_part2 <- cereals_df2[-split_index, ]

centroid1<- cereal_part1 %>% select_if(is.numeric) %>% filter(cluster==1) %>% colMeans()
centroid2 <- cereal_part1 %>% select_if(is.numeric) %>% filter(cluster==2) %>% colMeans()

centroid <- rbind(centroid1, centroid2)

cluster_B <- data.frame(data=seq(1,nrow(cereal_part2),1),clusterB = rep(0,nrow(cereal_part2)))

for(x in 1:nrow(cereal_part2))
{ cluster_B$clusterB <- which.min(as.matrix(get_dist(as.data.frame(rbind(centroid[,-25],cereal_part2[x,c(-1,-26)]))))[3,-3])}

cluster_B <- cluster_B %>% mutate(orig_clusters = cereal_part2$cluster)

mean(cluster_B$clusterB==cluster_B$orig_clusters)

```
Answer: The clusters are not stable as per the comparision.

Q4. The elementary public schools would like to choose a set of cereals to include in their daily cafeterias. Every day a different cereal is offered, but all cereals should support a healthy diet. For this goal, you are requested to find a cluster of “healthy cereals.” Should the data be normalized? If not, how should they be used in the cluster analysis?


```{r}

healthy_cereals <- data.frame(cereals_df2 %>% filter(cluster==2) %>% select_if(is.numeric) %>% colMeans())

```
Cluster 2 has cereals those are rich in protein, Vitamin, Carbs as well as low in sugar and sodium. Hence, Cereals in cluster 2 can be included to support a healthy diet

